---
title: "Covid91 vaccine study Final2023"
author: "Amber Walker!"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
FinalData$treatment <- relevel(FinalData$treatment,ref="placebo")
#FinalData$infected <- relevel(FinalData$infected,ref="OK")
males = filter(FinalData, sex =="M")
females = filter(FinalData, sex =="F")
LGBTQ = filter(FinalData, LGBTQ =="gay")
Druggies = filter(FinalData, DrugUser =="yes")
```

# Introduction

A new preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

#methods

The variables from FinalData needed to be changed from character variables into factor variables. Specific subgroups had to be separated from the broader categories in order to properly studied. Female, male, Druggies, and LGBTQ were all separated into subgroups. All of these categories are isolated to see weather or not these particular populations got sick and who did not from taking the placebo.

## Males

```{r}
dd2 <- males %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
There are two different plots based on males and their results of getting infected or not based on them taking the placebo or vaccine. The first plot will show side-by-side bars where the height of each bar represents the count of observations for each combination of infected and treatment. So it seems as though, there were very few males that received the placebo and vaccine that were infected and just a small percentage more were infected that received a placebo that a vaccine. In addition, a slightly more amount of males were not infected that received the vaccine and the placebo. The second plot will show side-by-side bars where the height of each bar represents the percentage of observations for each combination of infected and treatment. So it seems that more males were drastically infected and that was given the placebo over the vaccine. Whereas, the a slightly more amount of males that was given the placebo over the vaccine were not infected. 
 

### Numerical results

```{r}
table1 <- xtabs(~infected + treatment, data=males)
rowPerc(table1)
colPerc(table1)
```

These numerical results show specific percentages of males that got covid-19 after taking the placebo and taking the vaccine. 2.55 males were given the placebo and were infected, and 97.45 were not infected. This shows that many more people who were given the placebo were not infected as supposed to those who were infected. As far as the males that were given the vaccine, 1.15 were infected and 98.85 were not infected. This shows that there was a large number of males that were not infected after receiving the vaccine as supposed to those who did not. Based on this it seems that it was a much larger amount of males not being infected whether they took the vaccine or not. 

### Inferential 


```{r}
chisq.test(table1)
chisqtestGC(table1)
guysfish <- fisher.test(table1)
guysfish
```
Based on the results, both Pearson's Chi-squared test and Fisher's Exact Test indicate a significant association between the infected and treatment variables. The odds ratio indicates that the odds of being infected with covid19 are higher for individuals receiving the vaccine treatment compared to the placebo treatment, and this difference is statistically significant.

```{r}
effect <- guysfish$p.value/(guysfish$p.value +1)
effectivenessguy = guysfish$estimate/(guysfish$estimate +1)*100
print(effectivenessguy)
```

```{r}
low=(guysfish$conf.int[1]/(guysfish$conf.int[1] +1))*100
hi=(guysfish$conf.int[2]/(guysfish$conf.int[2] +1))*100
options(digits=2)
```


Above is the effectiveness for the covid-19 vaccine for males. Based on the interval seen above,the odds ratio for the  males are 69. The odds ratio is approximately how effective the vaccine is. We can conclude that the vaccine is effective in protecting guys from getting covid-19. 

## Females

```{r}
dd2 <- females %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
There are two different plots based on females and their results of getting infected or not based on them taking the placebo or vaccine. The first plot will show side-by-side bars where the height of each bar represents the count of observations for each combination of infected and treatment. So it seems as though, there were very few females that received the placebo and vaccine that were infected and just a small percentage more were infected that received a placebo that a vaccine. The same amount of females were not infected that received the vaccine and the placebo. The second plot will show side-by-side bars where the height of each bar represents the percentage of observations for each combination of infected and treatment.So it seems that many more females that were infected was given the placebo as supposed to the vaccine. However, the amount of females that were not infected was about the same amount of females that received the placebo and the vaccine. 

```{r}
table1 <- xtabs(~infected + treatment, data=females)
rowPerc(table1)
colPerc(table1)
``` 
The table provides shows that 60% of individuals infected with "covid19" received the placebo treatment, while 40% received the vaccine" treatment, and 50% of individuals classified as non infected received the "placebo" treatment, and 50% received the vaccine treatment.These numerical results show specific percentages of females that got covid-19 after taking the placebo and taking the vaccine. 2.0 females were given the placebo and were infected, and 98 were not infected.This shows that many more females who were given the placebo were not infected as supposed to those who were infected. As far as the females that were given the vaccine, 1.4 were infected and 98.6 were not infected. This shows that there was a large number of females that were not infected after receiving the vaccine as supposed to those who did not. Based on this it seems that it was a much larger amount of males not being infected whether they took the vaccine or not. 

## LGBTQ

## Druggies

# Overall Results and Conclusions
